### 2018-10-04 analysis/paper/SOM.Rmd
formatting to submit

---
title: "Supplementary Materials for 'The role of lithic bipolar technology in Western Iberia’s Upper Paleolithic: the case of Vale Boi (southern Portugal)'"
date: '`r Sys.Date()`'
output:
   bookdown::word_document2:
    fig_caption: no
    reference_docx: "../templates/template2.docx"
bibliography: ScaledPiecesVB.bib
csl: "../templates/springer-socpsych-author-date.csl"
---

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  dpi = 300,
  fig.width = 9,
  fig.path = "../figures/")
  cache = TRUE

library(readr)

```

\newline

### Table 1 - Technological attributes used in the analysis of scaled pieces included in this study

```{r techno_attributes, echo=FALSE}

techno_attributes <- read_csv("../data/raw_data/technological_attributes.csv")

techno_attributes[is.na(techno_attributes)] <- " "

knitr::kable(techno_attributes)

```

\newline

###  Table 2 - Morpho-functional attributes used to analyze each damaged platform of a scaled piece. *Adapted from Gonzalez-Urquijo and Ibanez-Estévez [-@gonzalez-urquijo_metodologialisis_1994]. **Adapted from de la Peña [-@de_la_pena_sobre_2011]

```{r morpho_attributes, echo=FALSE}

morpho_attributes <- read_csv("../data/raw_data/functional_attributes.csv")

morpho_attributes[is.na(morpho_attributes)] <- " "

knitr::kable(morpho_attributes)

```

\newline

# References 
<div id="refs"></div>

##### pagebreak

# Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```